---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  pdf_document: default
  html_document: default
  word_document: default
---
<center> <h1> Assignment 3 (10%) </h1> </center>
<center>  <h3> [Daria Yip] </h2> </center>
<center> <h3> [DHA - Student#500721106] </h2> </center>
---


#### 1. Import to R the following fiel: http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv (The dataset is related to white Portuguese "Vinho Verde" wine. For more info: https://archive.ics.uci.edu/ml/datasets/Wine+Quality) (3 points)
```{r}
library(RCurl) #getURL 
u <- getURL("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv")
whitewine <- read.csv(text = u, header = TRUE, sep = ";")
#This is a very important dataset to specify the separators,
#otherwise you will have a lot of observations in one column.
```

#### 2. Check the datatypes of the attributes. (3 points)
```{r}
str(whitewine)
#All the attributes are num type except quality which is int.
```
  
#### 3. Are there any missing values in the dataset? (4 points)
```{r}
sum(is.na(whitewine))
#There are no missing values in the dataset.
```

#### 4. What is the correlation between the attributes other than Quality? (10 points)
```{r}
correlation <- cor(whitewine[-12]) #want to have correlation of
#every attribute other than Quality which is variable 12
correlation
#Visualize the correlation
library(corrplot)
corrplot(correlation)
```

#### 5. Graph the frequency distribution of wine quality by using Quality. (10 points)
```{r}
hist(whitewine$quality, main = "Wine Quality", sub = "Wine Ratings")
```

#### 6.	Reduce the levels of rating for quality to three levels as high, medium and low. Assign the levels of 3 and 4 to level 0; 5 and 6 to level 1; and 7,8 and 9 to level 2. (10 points)
```{r}
whitewine$quality <- ifelse(whitewine$quality < 5, "Low", ifelse(whitewine$quality >= 7, "High", "Medium"))
head(whitewine)
```

#### 7. Normalize the data set by using the following function: (12 points)
```{r}
normalize <- function(x){
  return ((x - min(x)) / (max(x) - min(x)))
}
```

```{r}
whitewinen <- as.data.frame(lapply(whitewine[1:11], normalize))
#normalize the dataframe with lapply and only variables 1:11
summary(whitewine[1])
#checking the first variable to see if the data has been 
#normalized. We can see it has by the 0-1 scale.
```

#### 8. Divide the dataset to training and test sets. (12 points)
```{r}
trainingtesting <- sample(nrow(whitewine), floor(nrow(whitewine) * 0.7)) #70% is the training set and we will choose them 
#randomly with the "sample" function, floor will return a 
#numeric vector with the largest integers
training <- whitewinen[trainingtesting,]
test <- whitewinen[-trainingtesting,] #testing will be that #which is not the training
```

#### 9. Use the KNN algorithm to predict the quality of wine using its attributes. (12 points)
```{r}
#install.packages("class")
library(class)
training_labels <- whitewine[trainingtesting, 12]
test_labels <- whitewine[-trainingtesting, 12]
prediction <- knn(train = training, test = test, cl = training_labels, k = 10)
table(prediction)
```

#### 10. Display the confusion matrix to evaluate the model performance. (12 points)
```{r}
confusionmatrix <- table(Actual = test_labels, Predicted = prediction)
confusionmatrix
```

#### 11. Evaluate the model performance by computing Accuracy, Sensitivity and Specificity. (12 points)
```{r}
# install.packages("class")
#install.packages("gmodels")
library(gmodels)
#CrossTable(x = test_labels, y = prediction, prop.chisq = F)
accuracy <- mean(test_labels == prediction)
accuracy

error <- mean(test_labels != prediction)
error

#install.packages("caret", dependencies = TRUE)
library(caret)

test_labels <- as.factor(test_labels) #to make the levels
#the same for the confusionMatrix function

confusionMatrix(prediction, test_labels, positive = "pos")
#The model below shows the stats of accuracy which is 78.91%.
#Sensitivity = TruePositives/(TruePositives + FalseNegatives)
#Specificity is the True Negatives proportion.
#Specificity = TrueNegatives/(TrueNegatives + FalseNegatives)
#Precision = TruePositives/(TruePositives + FalsePositives).
```

This is the end of Assignment 3

Ceni Babaoglu, PhD