---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  word_document: default
  pdf_document: default
  html_document: default
---
<center> <h1> Assignment 1 (10%) </h1> </center>
<center>  <h3> [Daria Yip] </h2> </center>
<center> <h3> [DHA - Student#500721106] </h2> </center>
---

#### 1. Read the csv files in the folder. (3 points)
```{r}
csvdata1 <- read.csv("C:/Users/samda/Documents/Ryerson/CIND 642 - R/Macronutrients1.csv", header = T, sep =",")
csvdata2 <- read.csv("C:/Users/samda/Documents/Ryerson/CIND 642 - R/USDA_Micronutrients2.csv", header = T, sep =",")
```

#### 2. Merge the data frames using the variable "ID". Name the Merged Data Frame "USDA". (6 points)
```{r}
USDA <- merge(csvdata1, csvdata2, by = "ID") #merge function to merge two dataframes by a common variable

head(USDA)
```

#### 3. Check the datatypes of the attributes. Delete the commas in the Sodium and Potasium records. Assign Sodium and Potasium as numeric data types. (6 points)
```{r}
str(USDA) #Check the datatypes of the attributes, you can also use sapply(USDA, class)

USDA$Sodium <- as.numeric(gsub(",","",USDA$Sodium)) #delete commas in Sodium and assign
USDA$Potassium <- as.numeric(gsub(",","", USDA$Potassium)) #delete commas in Potassium & #assign to numeric

class(USDA$Sodium) #check class 
class(USDA$Potassium) #check class

head(USDA) #check that commas have been removed from both
```

#### 4. Remove records (rows) with missing values in more than 4 attributes (columns). How many records remain in the data frame? (6 points)
```{r}
missing <- rowSums(is.na(USDA))
USDA <- USDA[!missing > 4,]
nrow(USDA)

#Therefore, there are 6887 records that remain in the dataframe #that do not have rows with missing values more than 4 #attributes.
```

#### 5. For records with missing values for Sugar, Vitamin E and Vitamin D, replace missing values with mean value for the respective variable. (6 points)
```{r}
mean.sugar <- mean(USDA$Sugar, na.rm = TRUE)
USDA$Sugar[is.na(USDA$Sugar)] = mean.sugar

mean.vitd <- mean(USDA$VitaminD, na.rm=TRUE)
USDA$VitaminD[is.na(USDA$VitaminD)] = mean.vitd

mean.vite <- mean(USDA$VitaminE, na.rm=TRUE)
USDA$VitaminE[is.na(USDA$VitaminE)] = mean.vite

check <- sum(is.na(USDA$Sugar)) + sum(is.na(USDA$VitaminE)) + sum(is.na(USDA$VitaminD)) #double check #no more missing values in these columns
check
```

#### 6. With a single line of code, remove all remaining records with missing values. Name the new Data Frame "USDAclean". How many records remain in the data frame? (6 points)
```{r}
USDAclean <- na.omit(USDA) #single line of code to remove all remaining missing values
nrow(USDAclean) #remaining is 6310 from the original 6887 in Question 4
```

#### 7. Which food has the highest sodium level? (6 points)
```{r}
USDAclean[which(USDAclean$Sodium == max(USDAclean$Sodium)),]
#can also use the which.max #USDAclean[which.max(USDAclean$Sodium),]

#Therefore, salt has the highest sodium level.
```

#### 8. Create a histogram of Vitamin C distribution in foods. (6 points)
```{r}
summary(USDAclean$VitaminC) #take a look at the column to see
#what kind of axis limits we need 
hist(USDAclean$VitaminC, breaks = 100, xlim = c(0, 100),
     xlab = "Vitamin C", col = "yellow", main = "Vitamin C Distribution in Foods")
#breaks is very important to include here so R knows how we
#want the histogram to be broken up
```

#### 9. Create a boxplot to illustrate the distribution of values for TotalFat, Protein and Carbohydrate. (6 points)
```{r}
boxplot(USDAclean$TotalFat, USDAclean$Protein, USDAclean$Carbohydrate, names = c("Total Fat", "Protein", "Carbohydrate"), col = c("orange", "blue", "pink"), xlab = "Amount", main = "Distribution of Values for Total Fat, Protein, and Carbohydrates", horizontal = FALSE)
```

#### 10. Create a scatterplot to illustrate the relationship between a food's TotalFat content and its Calorie content. (6 points)
```{r}
plot(USDAclean$Calories~USDAclean$TotalFat, main = "Relationship between Calories and Total Fat Content", xlab = "Calories", ylab = "Total Fat Content", col = "purple")
```

#### 11. Add a variable to the data frame that takes value 1 if the food has higher sodium than average, 0 otherwise. Call this variable HighSodium. Do the same for High Calories, High Protein, High Sugar, and High Fat. How many foods have both high sodium and high fat? (8 points)
```{r}
HighSodium <- 0 #Initialize
HighCalories <- 0
HighProtein <- 0
HighSugar <- 0
HighFat <- 0

USDAclean <- cbind(USDAclean, HighSodium) #add column
USDAclean <- cbind(USDAclean, HighCalories, HighProtein,
                   HighSugar, HighFat) #add columns

#For loop to calculate the if and else for High Sodium #condition
for(i in 1:length(USDAclean$Description))
{
    if(USDAclean$Sodium[i] > mean(USDAclean$Sodium))
       {
         USDAclean$HighSodium[i] = 1
       }
       else
         {
           USDAclean$HighSodium[i] = 0
         }
}

#High Calories, using extraction
USDAclean$HighCalories[USDAclean$Calories > mean(USDAclean$Calories)] <- 1 
USDAclean$HighCalories[USDAclean$Calories <= mean(USDAclean$Calories)] <- 0

#HighProtein, using if else statement instead of for loop
USDAclean$HighProtein <- ifelse(USDAclean$Protein > mean(USDAclean$Protein),1,0)

#HighSugar
USDAclean$HighSugar <- ifelse(USDAclean$Sugar > mean(USDAclean$Sugar),1,0)

#HighFat
USDAclean$HighFat <- ifelse(USDAclean$TotalFat > mean(USDAclean$TotalFat),1,0)

index <- USDAclean[which(USDAclean$HighFat == 1 & USDAclean$HighSodium == 1),]
nrow(index) #High Sodium & High Fat foods

#Therefore, there are 644 foods that have high sodium and high
#fat content.

head(USDAclean) #take a look at the data
```

#### 12. Calculate the average amount of iron, for high and low protein foods. (8 points)
```{r}
highavg <- mean(USDAclean$Iron[USDAclean$HighProtein == 1]) #high #protein foods average iron
highavg

lowavg <- mean(USDAclean$Iron[USDAclean$HighProtein == 0]) #low
#protein foods average iron
lowavg

ironmean <- cbind(highavg,lowavg)
colnames(ironmean) <- c("High Protein", "Low Protein")
rownames(ironmean) <- c("Average Iron")
ironmean

#The average amount of iron for high protein foods is 3.069541
#for High Protein Levels and 2.696634 for low protein foods.

#You can also do this using the aggregate function, #aggregate(USDAclean$Iron), list(USDAclean$HighProtein) where #the FUN = mean.
```

#### 13. Create a script for a "HealthCheck" program to detect unhealthy foods. Use the algorithm flowchart below as a basis for this script. (8 points)
```{r}
require(jpeg)
img<-readJPEG("HealthCheck.jpg")
plot(1:4, ty = 'n', ann = F, xaxt = 'n', yaxt = 'n')
rasterImage(img,1,1,4,4)
```
```{r}
HC <- function(x,y,z)
{
  ifelse(x == 1 & y == 1 & z == 1, "Fail", "Pass")
        }
HC(0,1,1) #Testing 
```

#### 14. Add a new variable called HealthCheck to the data frame using the output of the function. (8 points)
```{r}
HealthCheck <- 0 #Initialize
USDAclean <- cbind(USDAclean, HealthCheck) #add column

USDAclean$HealthCheck <- HC(USDAclean$HighSodium,USDAclean$HighSugar,USDAclean$HighFat)

head(USDAclean)
```

#### 15. How many foods in the USDAclean data frame fail the HealthCheck? (8 points)
```{r}
sum(USDAclean$HealthCheck == "Fail")
```

#### 16. Save your final data frame as "USDAclean_ [your last name]." (3 points)
```{r}
USDAclean_Yip <- USDAclean
```


This is the end of Assignment 1

Ceni Babaoglu, PhD
